\section*{Acknowledgements}

\epigraph{\textit{``I've had it with these motherf****g snakes on this motherf****g plane''}}{ --- \textit{Samuel Jackson (Neville Flynn)}, \textbf{S}nakes \textbf{O}n \textbf{A} \textbf{P}lane}

There are many people to thank who have helped me throughout my PhD, who I will mention briefly here.
Firstly my supervisors Graham and Chris, who have always been around to offer me consistently great advice for the last 3.5 years.
Then Matt Pitkin, David Keitel and many other members of the IGR and LIGO collaboration who have helped me with various problems.
I would also like to thank Stewart Boogert and others at Royal Holloway who helped get me to this point and continue to offer advice whenever I see them.
Then the many friends who have kept me nice and distracted from my work: Alex, Abi, Paul, Ashlea, Simon, Joe, Martin, Andrew, Jenny and many others (you know who you are).
Finally I would like to thank my family for keeping me grounded and supporting me with whatever I wanted to do.

\section*{Declaration}

The material presented in this thesis is the result of my own work under the
supervision of Prof. Graham Woan and Dr Chris Messenger in the Institute for
Gravitational Research at the University of Glasgow. It has not been submitted
for any other degree at the University of Glasgow, or at any other institution.
This work was conducted as part of the \gls{LVK} collaboration and therefore
was made possible with the interaction of many researchers within this
collaboration.

\begin{itemize}
	\item Chapter \ref{intro} and Chapter \ref{searchcw} is introductory material written by myself under the advice of Prof. Graham Woan and Dr Chris Messenger.
	\item Chapter \ref{soap} is my own work which has been published in \citep{bayley2019SOAPGeneralised}, and was written in collaboration with my supervisors.
	\item Chapter \ref{machine} has been written into a paper which is yet to be published, but will be submitted soon. This was written in collaboration with my supervisors and was predominantly my own work.
	\item Chapter \ref{par_est} is work which is planned to be a paper in the future when the work is complete, and is my own work done in collaboration with my supervisors.
	\item Chapter \ref{detchar} is again work which is planned to be a paper in the future when the work is complete, and is my own work done in collaboration with my supervisors.
\end{itemize}